### 2023-03-04 manuscript.tex
Merge pull request #316 from SORTEE-Github-Hackathon/eric-edits [ci skip]
This build is based on
https://github.com/SORTEE-Github-Hackathon/manuscript/commit/4a0cc6accf8643e836371ea6dcf75962146c372d.

This commit was created by the following CI build and job:
https://github.com/SORTEE-Github-Hackathon/manuscript/commit/4a0cc6accf8643e836371ea6dcf75962146c372d/checks
https://github.com/SORTEE-Github-Hackathon/manuscript/actions/runs/4332730335

This manuscript
(\href{https://SORTEE-Github-Hackathon.github.io/manuscript/v/290c4cd8316ffaa08ceafd6ac4cfcdfe09774653/}{permalink})
was automatically generated
from \href{https://github.com/SORTEE-Github-Hackathon/manuscript/tree/290c4cd8316ffaa08ceafd6ac4cfcdfe09774653}{SORTEE-Github-Hackathon/manuscript@290c4cd}
on March 4, 2023.

@@

  Department of Wildlife Ecology and Conservation, University of Florida, Gainesville, FL, USA

@@

Most scientists, including ecologists and evolutionary biologists, depend on computational tools for their research (\protect\hyperlink{ref-fJWFe93e}{Hannay et al., 2009}).
Researchers write and use software packages or write code (hereafter, code) to perform scientific tasks ranging from data management, data analysis, and study replication, to the application and the development of tools for hypothesis testing.
Maintaining code for scientific collaboration requires an efficient and well-documented work-flow (\protect\hyperlink{ref-1Kqna6l2}{J. M. Perkel, 2020}).
To facilitate this process, scientists have been adopting tools from information and systems technology, such as cloud-based services for documentation and version control {[}\emph{e.g.}, the Google Suite, the Microsoft Suite, DropBox, and GitHub (\protect\hyperlink{ref-10ghgV3S8}{J. Perkel, 2016}){]}.
Within the spectrum of cloud-based services for collaboration, GitHub is uniquely positioned (\protect\hyperlink{tbl:compare}{Table 1}) to benefit scientists because it is designed to store, track changes, and enable collaboration on computer code -- fundamental components of modern research.
Albeit, most researchers lack exposure to adequate programming and coding practices and thus dedicate valuable time and effort to self-teaching research-facilitating tools.
GitHub is one of the most used web-based platforms for computational version control and collaboration, and in this paper, we provide researchers in ecology and evolutionary biology (EEB) with practical workflows to use GitHub to facilitate their collaborative research and code management practices.

@@

Version control involves tracking the state of the files and directories which are stored in a ``repository'' (see \protect\hyperlink{definitions}{Box 1}).
A basic workflow using Git and GitHub is to: (i) create a remote repository that is synchronized with files and directories stored locally; (ii) modify these files, either locally or remotely; (iii) frequently ``commit'' (or record) changes to these files (see \protect\hyperlink{definitions}{Box 1}) along with a description of modifications; (iv) synchronize commits with GitHub (see ``push'' and ``pull'' in \protect\hyperlink{definitions}{Box 1}) so that the repository on the web and the local repositories are up-to-date.
Repositories, which contain files, their modifications, and the description of their changes can then be accessed by chosen collaborators or, whenever applicable, the public, who can easily download and synchronize them to their own computers (see ``clone'' in \protect\hyperlink{definitions}{Box 1}).
Commits act like snapshots, one can view or even revert the state of the project to any previous commit.
If the modified files are plain text, only the differences from the previous commit are recorded, allowing you to frequently commit changes without causing the size of the project to grow excessively.
This provides a safe and less cluttered alternative to frequently making full copies of documents at different points in their evolution (\emph{e.g.} \texttt{analysis.R}, \texttt{analysis\_v2.R}, \texttt{analysis\_FINAL.R}).
While focusing on technical details about the use of Git and GitHub is beyond the scope of this study, we recommend users explore available resources to become more familiar with version control features (but see Perez-Riverol et al. (\protect\hyperlink{ref-kEX5dgzK}{2016}), Blischak et al. (\protect\hyperlink{ref-PlcxShQU}{2016}), and Kalliamvakou et al. (\protect\hyperlink{ref-ndfO9H}{2015})).

The expansive GitHub user-community and numerous GitHub resources have boosted its popularity {[}Perez-Riverol et al. (\protect\hyperlink{ref-kEX5dgzK}{2016}); Bryan (\protect\hyperlink{ref-RVetqmsg}{2018}); \url{https://happygitwithr.com}; \url{https://ourcodingclub.github.io}{]}.
Nevertheless, although multiple articles have encouraged researchers in EEB to adopt GitHub as part of their research process (\protect\hyperlink{ref-3DKwn1sY}{Lowndes et al., 2017}; \protect\hyperlink{ref-10ghgV3S8}{J. Perkel, 2016}), its use is still not widespread.
First-time users without formal training in information technology may face steep learning curves because GitHub and its features have been centered on software development (\protect\hyperlink{ref-139b0pSGc}{Leibzon, 2016}).
Moreover, domain-specific resources providing tractable examples and practical guidance for researchers in EEB on GitHub are scarce {[}but see Kim et al. (\protect\hyperlink{ref-lJAgyhYq}{2022}); \url{https://ourcodingclub.github.io}; \url{https://www.openscapes.org}{]}.
Widespread adoption of GitHub for collaborating on research tasks can ultimately enable EEB researchers to spend less time on creating novel processes for collaboration and more time on their research (\protect\hyperlink{ref-ydrk01SR}{Briney et al., 2020}).
More importantly, expanding the availability of data and code management standards---of which GitHub is one increasingly important component---makes research more reproducible and collaborative (\protect\hyperlink{ref-13QX8XU3J}{Alston \& Rick, 2021}; \protect\hyperlink{ref-pq2Tv1BC}{Gomes et al., 2022b}).

@@

\hypertarget{storing-and-sharing-research-compendia}{%
\subsubsection{Storing and sharing research compendia}\label{storing-and-sharing-research-compendia}}

An EEB research compendium includes all computational materials related to research production, including data, code for analyses and protocols.
Safely storing these files is essential to protect against accidental modifications or deletions.
Many researchers begin using GitHub to backup their research compendium (\protect\hyperlink{ref-MwwMapRG}{Marwick et al., 2018}) into a centralized, readily-available remote server (see \protect\hyperlink{definitions}{Box 1}).
This practice has the advantages of facilitating collaboration, integrating data and code archiving, allowing file versions to be accessed and restored, and further contributing to open science practices (Borghi \& Van Gulick (\protect\hyperlink{ref-gLby7jt1}{2022})).
Every change made to files in version-controlled repositories accompany authored descriptions of modifications (\protect\hyperlink{definitions}{Box 1}).
Later, the entire history of commits and their commit messages are viewable and can be audited similarly to physical laboratory notebooks (\protect\hyperlink{ref-4ny1onB0}{Ram, 2013}).
GitHub allows administrative control over who can view and modify repositories, and the process of cloning, forking and suggesting changes through pull requests (see \protect\hyperlink{definitions}{Box 1}) ensures code owners have full control over code and documents.
Repositories can be changed from public to private, allowing data storage and management without sacrificing the privacy that may be necessary for certain compendia.
With this, researchers can grant collaborators read and/or write access to files in private repositories to pursue pre-publication analyses or writing in privacy.

@@

A third barrier may come from general reluctance to share data and code publicly, or technical and logistical issues (\protect\hyperlink{ref-pq2Tv1BC}{Gomes et al., 2022b}).
GitHub is, by default, a public and open platform.
This openness may add additional pressure to students and scientists learning to use the platform.
Moreover, additional tools may be required to fully integrate project files and GitHub repositories (\emph{e.g.}, \protect\hyperlink{ref-lx49NGto}{\emph{Connect GitHub to a Project - OSF Support}, n.d.}).
Other scientists may simply lack the time or incentives to document and version control their code if the code is unlikely to be reused beyond their analysis.
However we (and others, \emph{e.g.,} \protect\hyperlink{ref-VDJput1V}{Gomes et al., 2022a}) argue that the endeavour of open science and collaboration requires code owners to document and version control code despite uncertainty around future use.

@@

\leavevmode\vadjust pre{\hypertarget{ref-VDJput1V}{}}%
Gomes, D. G. E., Pottier, P., Crystal-Ornelas, R., Hudgins, E. J., Foroughirad, V., Sánchez-Reyes, L. L., Turba, R., Martinez, P. A., Moreau, D., Bertram, M. G., Smout, C. A., \& Gaynor, K. M. (2022a). Why don't we share data and code? Perceived barriers and benefits to public archiving practices. \emph{Proceedings of the Royal Society B: Biological Sciences}, \emph{289}(1987). \url{https://doi.org/10.1098/rspb.2022.1113}

\leavevmode\vadjust pre{\hypertarget{ref-pq2Tv1BC}{}}%
Gomes, D. G. E., Pottier, P., Crystal-Ornelas, R., Hudgins, E. J., Foroughirad, V., Sánchez-Reyes, L. L., Turba, R., Martinez, P. A., Moreau, D., Bertram, M., Smout, C., \& Gaynor, K. (2022b). \emph{Why don't we share data and code? Perceived barriers and benefits to public archiving practices}. Center for Open Science. \url{https://doi.org/10.31222/osf.io/gaj43}

>>>>

This manuscript
(\href{https://SORTEE-Github-Hackathon.github.io/manuscript/v/4a0cc6accf8643e836371ea6dcf75962146c372d/}{permalink})
was automatically generated
from \href{https://github.com/SORTEE-Github-Hackathon/manuscript/tree/4a0cc6accf8643e836371ea6dcf75962146c372d}{SORTEE-Github-Hackathon/manuscript@4a0cc6a}
on March 4, 2023.

@@

  Communications \& Cyber Techologies, University of Arizona, Tucson, USA

@@

Most scientists, including ecologists and evolutionary biologists, depend on computational tools for their research (\protect\hyperlink{ref-fJWFe93e}{Hannay et al., 2009}).
Researchers write and use software packages or write code (hereafter, code) to perform scientific tasks ranging from data management, data analysis, and study replication, to the application and the development of tools for hypothesis testing.
Maintaining code for scientific collaboration requires an efficient and well-documented work-flow (\protect\hyperlink{ref-1Kqna6l2}{J. M. Perkel, 2020}).
To facilitate this process, scientists have been adopting tools from information and systems technology, such as cloud-based services for documentation and version control {[}\emph{e.g.}, the Google Suite, the Microsoft Suite, DropBox, and GitHub (\protect\hyperlink{ref-10ghgV3S8}{J. Perkel, 2016}){]}.
Within the spectrum of cloud-based services for collaboration, GitHub is uniquely positioned (\protect\hyperlink{tbl:compare}{Table 1}) to benefit scientists because it is designed to store, track changes, and enable collaboration on computer code---fundamental components of modern research.
Albeit, most researchers lack exposure to adequate programming and coding practices and thus dedicate valuable time and effort to self-teaching research-facilitating tools.
GitHub is one of the most used web-based platforms for computational version control and collaboration, and in this paper, we provide researchers in ecology and evolutionary biology (EEB) with practical workflows to use GitHub to facilitate their collaborative research and code management practices.

@@

Version control involves tracking the state of the files and directories which are stored in a ``repository'' (see \protect\hyperlink{definitions}{Box 1}).
A basic workflow using Git and GitHub is to: (i) create a remote repository that is synchronized with files and directories stored locally; (ii) modify these files, either locally or remotely; (iii) frequently ``commit'' (or record) changes to these files (see \protect\hyperlink{definitions}{Box 1}) along with a description of modifications; (iv) synchronize commits with GitHub (see ``push'' and ``pull'' in \protect\hyperlink{definitions}{Box 1}) so that the repository on the web and the local repositories are up-to-date.
Repositories, which contain files, their modifications, and the description of their changes can then be accessed by chosen collaborators or, whenever applicable, the public, who can easily download and synchronize them to their own computers (see ``clone'' in \protect\hyperlink{definitions}{Box 1}).
Commits act like snapshots, one can view or even revert the state of the project to any previous commit.
If the modified files are plain text, only the differences from the previous commit are recorded, allowing you to frequently commit changes without causing the size of the project to grow excessively.
This provides a safe and less cluttered alternative to frequently making full copies of documents at different points in their evolution (\emph{e.g.} \texttt{analysis.R}, \texttt{analysis\_v2.R}, \texttt{analysis\_FINAL.R}).
While focusing on technical details about the use of Git and GitHub is beyond the scope of this study, we recommend users explore available resources to become more familiar with version control features (see Perez-Riverol et al. (\protect\hyperlink{ref-kEX5dgzK}{2016}), Blischak et al. (\protect\hyperlink{ref-PlcxShQU}{2016}), and Kalliamvakou et al. (\protect\hyperlink{ref-ndfO9H}{2015})).

The expansive GitHub user-community and numerous GitHub resources have boosted its popularity {[}Perez-Riverol et al. (\protect\hyperlink{ref-kEX5dgzK}{2016}); Bryan (\protect\hyperlink{ref-RVetqmsg}{2018}); \url{https://happygitwithr.com}; \url{https://ourcodingclub.github.io}{]}.
Nevertheless, although multiple articles have encouraged researchers in EEB to adopt GitHub as part of their research process (\protect\hyperlink{ref-3DKwn1sY}{Lowndes et al., 2017}; \protect\hyperlink{ref-10ghgV3S8}{J. Perkel, 2016}), its use is still not widespread.
First-time users without formal training in information technology may face steep learning curves because GitHub and its features have been centered on software development (\protect\hyperlink{ref-139b0pSGc}{Leibzon, 2016}).
Moreover, domain-specific resources providing tractable examples and practical guidance for researchers in EEB on GitHub are scarce {[}but see Kim et al. (\protect\hyperlink{ref-lJAgyhYq}{2022}); \url{https://ourcodingclub.github.io}; \url{https://www.openscapes.org}{]}.
Widespread adoption of GitHub for collaborating on research tasks can ultimately enable EEB researchers to spend less time on creating novel processes for collaboration and more time on their research (\protect\hyperlink{ref-ydrk01SR}{Briney et al., 2020}).
More importantly, expanding the availability of data and code management standards---of which GitHub is one increasingly important component---makes research more reproducible and collaborative (\protect\hyperlink{ref-13QX8XU3J}{Alston \& Rick, 2021}; \protect\hyperlink{ref-VDJput1V}{Gomes et al., 2022}).

@@

\hypertarget{storing-sharing}{%
\subsubsection{Storing and sharing research compendia}\label{storing-sharing}}

An EEB research compendium includes all computational materials related to research production, including data, code for analyses and protocols.
Safely storing these files is essential to protect against accidental modifications or deletions.
Many researchers begin using GitHub to backup their research compendium (\protect\hyperlink{ref-MwwMapRG}{Marwick et al., 2018}) into a centralized, readily-available remote server (see \protect\hyperlink{definitions}{Box 1}).
This practice has the advantages of facilitating collaboration, integrating data and code archiving, allowing file versions to be accessed and restored, and further contributing to open science practices (\protect\hyperlink{ref-gLby7jt1}{Borghi \& Van Gulick, 2022}).
Changes made to files in version-controlled repositories are accompanied by authored descriptions of modifications (\protect\hyperlink{definitions}{Box 1}).
Later, the entire history of commits and their commit messages are viewable and can be audited similarly to physical laboratory notebooks (\protect\hyperlink{ref-4ny1onB0}{Ram, 2013}).
GitHub allows administrative control over who can view and modify repositories, and the process of cloning, forking and suggesting changes through pull requests (see \protect\hyperlink{definitions}{Box 1}) ensures code owners have full control over code and documents.
Repositories can be changed from public to private, allowing data storage and management without sacrificing the privacy that may be necessary for certain compendia.
With this, researchers can grant collaborators read and/or write access to files in private repositories to pursue pre-publication analyses or writing in privacy.

@@

A third barrier may come from general reluctance to share data and code publicly, or technical and logistical issues (\protect\hyperlink{ref-VDJput1V}{Gomes et al., 2022}).
GitHub is, by default, a public and open platform.
This openness may add additional pressure to students and scientists learning to use the platform.
Moreover, additional tools may be required to fully integrate project files and GitHub repositories (\emph{e.g.}, \protect\hyperlink{ref-lx49NGto}{\emph{Connect GitHub to a Project - OSF Support}, n.d.}).
Other scientists may simply lack the time or incentives to document and version control their code if the code is unlikely to be reused beyond their analysis.
However we (and others, \emph{e.g.,} \protect\hyperlink{ref-VDJput1V}{Gomes et al., 2022}) argue that the endeavour of open science and collaboration requires code owners to document and version control code despite uncertainty around future use.

@@

\leavevmode\vadjust pre{\hypertarget{ref-VDJput1V}{}}%
Gomes, D. G. E., Pottier, P., Crystal-Ornelas, R., Hudgins, E. J., Foroughirad, V., Sánchez-Reyes, L. L., Turba, R., Martinez, P. A., Moreau, D., Bertram, M. G., Smout, C. A., \& Gaynor, K. M. (2022). Why don't we share data and code? Perceived barriers and benefits to public archiving practices. \emph{Proceedings of the Royal Society B: Biological Sciences}, \emph{289}(1987). \url{https://doi.org/10.1098/rspb.2022.1113}
